(n,p) Cross-Section Measurements for \ce{^{64}Cu} and \ce{^{47}Sc} via D-D Neutron Generator\\ OR \\(n,p) Cross-Section Measurements for \ce{^{64}Cu} and \ce{^{47}Sc} at 2.76 MeV using a D-D Neutron Generator for Medical Isotope Studies

### 2016-12-22 First_np_Paper/first_np_paper.tex
Updated introduction section with first-round edits

A series of experiments are being conducted at UC Berkeley to measure low-energy neutron-induced production cross sections for a range of emerging medical radioisotopes. One potential method for producing many of these isotopes is the use of the (n,p) reaction via D-D neutron generators. However, many of these cross sections currently have significant uncertainties. Direct neutron activation, using a D-D neutron generator, allows for more precise cross section measurements than time-of-flight methods, and potentially offers a proliferation-resistant pathway for radioisotope production. Neutron activation can provide several advantages over charged-particle reactions from a targetry perspective. Notably, the higher penetration of neutrons allows for the use of thick target designs, which can improve production throughput and greatly simplifies target cooling considerations. More importantly, low-energy neutron reactions can be used to prevent competing reaction channels in a charged-particle reaction which would otherwise contaminate the product specific activity.

The present work focuses on the (n,p) production cross section measurements of \ce{^{64}Cu} and \ce{^{47}Sc}. \ce{^{64}Cu} ($t_{1/2}$ = 12.7 h) undergoes $\beta^+$ decay (with a 61.5\% branching ratio) to \ce{^{64}Ni} or $\beta^-$ decay (with a 38.5\% branching ratio) to \ce{^{64}Zn}. The emitted 190-keV $\beta^-$ particle makes this an extremely attractive short-ranged therapeutic radionuclide, which also has the possibility for simultaneous PET imaging due to the positron emission branch. This combination of simultaneous imaging (for real-time dose monitoring and verification) and therapeutic properties makes \ce{^{64}Cu} particularly attractive for emerging radiation therapy protocols. In addition, copper radiochemistry is well-developed, and many existing ligands and carriers may be used for selective delivery of the radionuclide to different sites in patients.

\ce{^{47}Sc} ($t_{1/2}$ = 3.35 d) undergoes $\beta^-$ decay to \ce{^{47}Ti}, emitting a high-intensity 159-keV gamma ray in the process. This radionuclide is particularly attractive as an emerging diagnostic isotope, due to the similarity of the emitted gamma ray to that of the more well-established \ce{^{99m}Tc}. However, due to the short half-life  and dwindling supplies of \ce{^{99m}Tc}, \ce{^{47}Sc} stands poised as a potential solution to this shortage, due to its longer half-life and multiple production pathways without the need for highly enriched uranium.

Current methodology in radiochemistry has been proven to successfully recover upwards of 95\% of produced \ce{^{64}Cu} and \ce{^{47}Ti} from solid targetry designs, without the need for additional carrier. By expanding the base of efficient reaction pathways, great advances can be made in making production of medical radionuclides more efficient and affordable for those in need.

>>>>

%
% If we want to mention in-situ-production
%
%
% e.g.:In-situ production (I always figured this is what is so great, correct me if I am wrong) of radioactive nuclides is an emerging method for radiation therapy in nuclear medicine (e.g., cite some recent review paper). It is based on doping the malignant target tissue with a stable nuclide which is then activated in-situ by neutron irradiation. Following decay is focused on the target tissue. Advantages compared to conventional radiation therapy are the reduction of damage to surrounding cells, … short range, … . Currently development of this technique is hampered by the lack of fundamental cross section data for various isotopes.

\comment{Add an initial \enquote{punchline} to  clarify the motivation?}

\comment{Per Mauricio, does 2.7 MeV count as high energy? While fast, I would consider this to still be definitely in the low-energy regime, compared to spallation, etc.}

A series of experiments are being conducted at UC Berkeley to measure low-energy neutron-induced production cross sections for a range of emerging medical radioisotopes, as many of these cross sections currently have not been measured, or possess significant uncertainties.   One potential method for producing many of these isotopes is the use of (n,p) reactions via neutron beams from D-D neutron generators. Direct neutron activation, using a D-D neutron generator, allows for more precise cross section measurements than time-of-flight methods, and potentially offers a proliferation-resistant pathway for radioisotope production. Neutron activation can provide several advantages over charged-particle reactions from a targetry perspective. Notably, the higher penetration of neutrons allows for the use of thick target designs, which can improve production throughput and greatly simplifies target cooling considerations. More importantly, low-energy neutron reactions can be used to prevent competing reaction channels in a charged-particle reaction which would otherwise contaminate the product specific activity.

The present work focuses on the (n,p) production cross section measurements of \ce{^{64}Cu} and \ce{^{47}Sc}. Copper-64  ($t_{1/2}$ = 12.7 h) undergoes $\beta^+$ decay (61.5\% branching ratio) to \ce{^{64}Ni} or $\beta^-$ decay (38.5\% branching ratio) to \ce{^{64}Zn}. The emitted 190-keV $\beta^-$ particle makes this an  attractive short-ranged therapeutic radionuclide, which also has the possibility for simultaneous positron emission tomography (PET) imaging  due to the positron emission branch (for real-time dose monitoring and verification),  making \ce{^{64}Cu} particularly attractive for emerging radiation therapy protocols. In addition, copper radiochemistry is well developed, and many existing ligands and carriers may be used for selective delivery of the radionuclide to different sites in patients.

Scandium-47 ($t_{1/2}$ = 3.35 d) undergoes $\beta^-$ decay to \ce{^{47}Ti}, emitting a high-intensity 159-keV gamma ray in the process. This radionuclide is  attractive as an emerging diagnostic isotope, due to the similarity of the emitted gamma ray to that of the more well-established \ce{^{99m}Tc}. However, due to the short half-life  and dwindling supplies of \ce{^{99m}Tc}, \ce{^{47}Sc} stands poised as a potential solution to this shortage, due to its longer half-life and multiple production pathways without the need for highly enriched uranium.

\comment{Adding a short paragraph from Mauricio explaining benefits of DD generators vs other neutron sources?}

\comment{Per Daniel, add a short summary of results here?}

Current methodology in radiochemistry has been proven to successfully recover upwards of 95\% of produced \ce{^{64}Cu} and \ce{^{47}Ti} from solid target designs, without the need for additional carrier. By expanding the base of efficient reaction pathways, great advances can be made in making production of medical radionuclides more efficient and affordable for those in need.

\comment{Should this last paragraph be moved to the results and discussion at the end of the paper?}
